\label{Sec:Biblio}
\begin{thebibliography}{1}
\bibitem{AllCah79}
S.M. Allen, and J.W. Cahn,
\newblock A microscopic theory for antiphase boundary motion and its applications to antiphase domain coarsening,
\newblock Acta Metallurgica \textbf{27}, 1085-1095, (1979).
\bibitem{BirRot89} 
G. Birkhoff, and G.\textendash{}C., Rota,  
\newblock \textbf{Ordinary Differential Equations} (4th ed.), 
\newblock Wiley,  (1989).
\bibitem{BlaCasChaMur12}
S. Blanes, F. Casas, P. Chartier and A. Murua,
\newblock Splitting methods with complex coefficients for some classes of evolution equations,
\newblock Mathematics of Computation (forthcoming)
\newblock \url{http://arxiv.org/abs/1102.1622}
\bibitem{Bra06}
B. Bradie,
\newblock \textbf{A Friendly Introduction to Numerical Analysis},
\newblock Pearson, (2006).
\bibitem{BofEck12}
G. Boffetta and R.E. Ecke,
\newblock Two-Dimensional Turbulence,
\newblock Annual Review of Fluid Mechanics \textbf{44}, 427-451, (2012).
\bibitem{BoyDip10}
W.E. Boyce and R.C. DiPrima,
\newblock \textbf{Elementary Differential Equations and Boundary Value Problems},
\newblock Wiley, (2010).
\bibitem{Boy01}
J. P. Boyd,
\newblock \textbf{Chebyshev and Fourier Spectral Methods},
\newblock Dover, (2001).
\newblock \url{http://www-personal.umich.edu/~jpboyd/}
\bibitem{CHQZ06}
C. Canuto, M.Y. Hussaini, A. Quarteroni and T.A. Zang,
\newblock \textbf{Spectral Methods: Fundamentals in Single Domains},
\newblock Springer, (2006).
\bibitem{CHQZ07}
C. Canuto, M.Y. Hussaini, A. Quarteroni and T.A. Zang,
\newblock \textbf{Spectral Methods: Evolution to Complex Geometries and Applications to Fluid Dynamics},
\newblock Springer, (2007).
\bibitem{CicBra05}
C. Cichowlas and M.-E. Brachet,
\newblock Evolution of complex singularities in Kida-Pelz and Taylor-Green inviscid flows,
\newblock Fluid Dynamics Research \textbf{36}, 239-248, (2005).
\bibitem{CloMuiRig12}
\newblock Performance of FORTRAN and C GPU Extensions for a Benchmark Suite of Fourier Pseudospectral Algorithms
\newblock Forthcoming Proceedings of the Symposium on Application Accelerators in High Performancs computing (2012)
\newblock \url{http://arxiv.org/abs/1206.3215}
\bibitem{CooTuk65}
J.W. Cooley and J.W. Tukey,
\newblock An algorithm for the machine calculation of complex Fourier series,
\newblock Mathematics of Computation \textbf{19}, 297-301, (1965). 
\bibitem{CouJoh98}
R. Courant and F. John,
\newblock \textbf{Introduction to Calculus and Analysis I, II}
\newblock Springer (1998,1999)
\bibitem{DonSch11}
R. Donninger and W. Schlag,
\newblock Numerical study of the blowup/global existence dichotomy for the focusing cubic nonlinear Klein-Gordon equation,
\newblock Nonlinearity \textbf{24}, 2547-2562, (2011).
\bibitem{DoeGib95}
C.R. Doering and J.D. Gibbon,
\newblock \textbf{Applied Analysis of the Navier-Stokes Equations},
\newblock Cambridge University Press, (1995).
\bibitem{EliShu09}
B. Eliasson and P. K. Shukla
\newblock Nonlinear aspects of quantum plasma physics: Nanoplasmonics and nanostructures in dense plasmas
\newblock Plasma and Fusion Research: Review Articles, \textbf{4}, 32 (2009). 
\bibitem{Eva10}
L.C. Evans,
\newblock \textbf{Partial Differential Equations},
\newblock American Mathematical Society, (2010).
\bibitem{For77}
B. Fornberg,
\newblock A numerical study of 2-D turbulence,
\newblock Journal of Computational Physics \textbf{25}, 1-31, (1977).
\bibitem{Gal02}
G. Gallavotti,
\newblock \textbf{Foundations of Fluid Dynamics},
\newblock Springer, (2002).\\
\newblock \url{http://www.math.rutgers.edu/~giovanni/glib.html#E}
\bibitem{GotOrs77}
D. Gottlieb and S.A. Orszag,
\newblock \textbf{Numerical Analysis of Spectral Methods: Theory and Applications},
\newblock SIAM, (1977).
\bibitem{Gre94}
W. Grenier,
\newblock \textbf{Relativistic Quantum Mechanics},
\newblock Springer, (1994)
\bibitem{GroLusSkj99}
W. Gropp, E. Lusk and A. Skjellum,
\newblock \textbf{Using MPI},
\newblock MIT Press, (1999).
\bibitem{GroLusTha99}
W. Gropp, E. Lusk and R. Thakur,
\newblock \textbf{Using MPI-2},
\newblock MIT Press, (1999).
\bibitem{HeiJohBur84}
M.T. Heideman, D.H. Johnson and C.S. Burrus,
\newblock Gauss and the History of the Fast Fourier Transform,
\newblock  IEEE ASSP Magazine \textbf{1}(4), 14Ð21, (1984).
\bibitem{HesGotGot07}
J.S. Hesthaven, S. Gottlieb and D. Gottlieb, 
\newblock \textbf{Spectral Methods for Time-Dependent Problems},
\newblock Cambridge University Press, (2007).
\bibitem{HugEtAl08}
D. Hughes-Hallett, A.M. Gleason, D.E. Flath, P.F. Lock, D.O. Lomen, D. Lovelock, W.G. MacCallum, D. Mumford, B. G. Osgood, D. Quinney, K. Rhea, J. Tecosky-Feldman, T.W. Tucker, and O.K. Bretscher, A. Iovita, W. Raskind, S.P. Gordon, A. Pasquale, J.B. Thrash,
\newblock \textbf{Calculus, Single and Multivariable}, 5th ed.
\newblock Wiley, (2008)
\bibitem{HolKarLieRis10}
H. Holden, K.H. Karlsen, K.-A. Lie and N.H. Risebro,
\newblock \textbf{Splitting Methods for Partial Differential Equations with Rough Solutions},
\newblock European Mathematical Society Publishing House, Zurich, (2010).
\bibitem{HolKarRisTao11}
H. Holden, K.H. Karlsen, N.H. Risebro and T. Tao,
\newblock Operator splitting for the KdV equation,
\newblock Mathematics of Computation \textbf{80}, 821-846, (2011).
\bibitem{Ise09}
 A. Iserles,  
\newblock \textbf{A First Course in the Numerical Analysis of Differential Equations}, 
\newblock Cambridge University Press, (2009).
\bibitem{Joh12}
R. Johnstone,
\newblock Improved Scaling for Direct Numerical Simulations of Turbulence,
\newblock HECTOR distributed Computational Science and Engineering Report,
\newblock \url{http://www.hector.ac.uk/cse/distributedcse/reports/ss3f-swt/}
\bibitem{Kle08}
C. Klein,
\newblock Fourth order time-stepping for low dispersion Korteweg-De Vries and nonlinear Schr\"{o}dinger equations,
\newblock Electronic Transactions on Numerical Analysis \textbf{29}, 116-135, (2008).
\bibitem{KleMuiRoi11}
C. Klein, B.K. Muite and K. Roidot,
\newblock Numerical Study of Blowup in the Davey-Stewartson System, 
\newblock \url{http://arxiv.org/abs/1112.4043}
\bibitem{KleRoi11}
C. Klein and K. Roidot,
\newblock Fourth order time-stepping for Kadomstev-Petviashvili and Davey-Stewartson Equations,
\newblock SIAM Journal on Scientific Computation \textbf{33}, 3333-3356, (2011).\\
\newblock \url{http://arxiv.org/abs/1108.3345}
\bibitem{LaiLam09}
S. Laizet and E. Lamballais,
\newblock High-order compact schemes for incompressible flows: A simple and efficient method with quasi-spectral accuracy,
\newblock Journal of Computational Physics \textbf{228}, 5989-6015, (2009).
\bibitem{LaiLi11}
S. Laizet and N. Li,
\newblock Incompact3d: A powerful tool to tackle turbulence problems with up to $O(10^5)$ computational cores,
\newblock International Journal of Numerical Methods in Fluids \textbf{67}, 1735-1757, (2011).
\bibitem{Lan96}
R. H. Landau,
\newblock \textbf{Quantum Mechanics II},
\newblock Wiley, (1996).
\bibitem{LaxBurLax76}
P. Lax, S. Burstein and A. Lax,
\newblock \textbf{Calculus with Applications and Computing, Vol. 1},
\newblock Springer, (1976).
\bibitem{LiLai10}
N. Li and S. Laizet,
\newblock 2DECOMP\&FFT - A highly scalable 2D decomposition library and FFT interface,
\newblock Proc. Cray User Group 2010 Conference.\\
\newblock \url{http://www.2decomp.org/pdf/17B-CUG2010-paper-Ning_LI.pdf}
\bibitem{LieLos03}
E.H. Lieb and  M. Loss,
\newblock \textbf{Analysis},
\newblock American Mathematical Society, (2003).
\bibitem{LinPon09}
F. Linares and G. Ponce,
\newblock \textbf{Introduction to Nonlinear Dispersive Equations},
\newblock Springer, (2009).
\bibitem{LevWag11}
J. Levesque and G. Wagenbreth,
\newblock \textbf{High Performance Computing: Programming and Applications},
\newblock CRC Press, (2011).
\bibitem{MajBer02}
A.J. Majda and A.L. Bertozzi,
\newblock \textbf{Vorticity and Incompressible Flow},
\newblock Cambridge University Press, (2002).
\bibitem{McLQui02}
R.I. McLachlan and G.R.W. Quispel,
\newblock Splitting Methods,
\newblock Acta Numerica \textbf{11}, 341-434, (2002).
\bibitem{MetReiCoh11}
M. Metcalf, J. Reid and M. Cohen,
\newblock \textbf{Modern Fortran Explained},
\newblock Oxford University Press, (2011).
\bibitem{NakSch11}
K. Nakanishi and W. Schlag,
\newblock \textbf{Invariant Manifolds and Dispersive Hamiltonian Evolution Equations},
\newblock European Mathematical Society, (2011).
\bibitem{Olv10}
P.J. Olver,
\newblock Dispersive Quantization,
\newblock American Mathematical Monthly, \textbf{117}, 599-610, (2010).
\bibitem{OlvSha06}
P.J. Olver and C. Shakiban,
\newblock \textbf{Applied Linear Algebra},
\newblock  Prentice Hall, (2006).
\bibitem{OrsPat72}
S.A. Orszag and G.S. Patterson Jr.,
\newblock Numerical simulation of three-dimensional homogeneous isotropic turbulence,
\newblock Physical Review Letters  \textbf{28(2)}, 76-79, (1972).
\bibitem{Pey02}
R. Peyret,
\newblock \textbf{Spectral Methods for Incompressible Viscous Flow},
\newblock Springer, (2002).
\bibitem{RenRog04}
R. Renardy and R.C. Rogers,
\newblock \textbf{An Introduction to Partial Differential Equations},
\newblock Springer, (2004).
\bibitem{Sha93}
A. Shapiro
\newblock The use of an exact solution of the Navier-Stokes equations in a validation test of a three-dimensional nonhydrostatic numerical model,
\newblock Monthly Weather Review \textbf{121}, 2420-2425, (1993).
\bibitem{SheTanWan11}
J. Shen, T. Tang and L.-L. Wang,
\newblock \textbf{Spectral Methods: Algorithms, Analysis and Applications},
\newblock Springer, (2011).
\bibitem{SulSul99}
C. Sulem and P.L. Sulem,
\newblock \textbf{The Nonlinear Schr\"{o}dinger equation: Self-Focusing and Wave Collapse},
\newblock Springer, (1999).
\bibitem{Tem01}
R. Temam,
\newblock \textbf{Navier-Stokes Equations},
\newblock Third revised edition, AMS, (2001).
\bibitem{Tha08}
M. Thalhammer,
\newblock Time-Splitting Spectral Methods for Nonlinear Schr\"{o}dinger Equations,
\newblock Unpublished manuscript, (2008).\\
\newblock \url{http://techmath.uibk.ac.at/mecht/research/SpringSchool/manuscript_Thalhammer.pdf}
\bibitem{Tre00}
L. N. Trefethen,
\newblock \textbf{Spectral Methods in Matlab},
\newblock SIAM, (2000).
\bibitem{TreEmb01}
\newblock \textbf{The (Unfninished) PDE coffee table book.}
L. N. Trefethen and K. Embree (Ed.),
\newblock Unpublished notes available online\\
\newblock \url{http://people.maths.ox.ac.uk/trefethen/pdectb.html} 
\bibitem{Tri88}
D.J. Tritton,
\newblock \textbf{Physical Fluid Dynamics},
\newblock Clarendon Press, (1988).
\bibitem{Uec09}
H. Uecker,
\newblock A short ad hoc introduction to spectral for parabolic PDE and the Navier-Stokes equations,
\newblock Lecture notes from the 2009 International Summer School on Modern Computational Science\\
\newblock \url{http://www.staff.uni-oldenburg.de/hannes.uecker/hfweb-e.html}
\bibitem{WeiHer86}
J.A.C. Weideman and B.M. Herbst,
\newblock Split-step methods for the solution of the nonlinear Schr\"{o}dinger equation,
\newblock SIAM Journal on Numerical Analysis \textbf{23(3)}, 485-507, (1986).
\bibitem{Yan10}
J. Yang,
\newblock \textbf{Nonlinear Waves in Integrable and Nonintegrable Systems},
\newblock SIAM, (2010).
\bibitem{Yan06}
L. Yang,
\newblock Numerical studies of the Klein-Gordon-Schr\"{o}dinger equations,
\newblock Masters Thesis, National University of Singapore\\
\newblock \url{http://scholarbank.nus.edu.sg/handle/10635/15515}
\end{thebibliography}
